---
title: "Geography 13"
author: "[Hannah Malak](https://hannahmalak.github.io/)"
subtitle: 'Lab 02: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
---

```{r, echo=TRUE, include=FALSE}
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)
```

# Data

```{r, message = FALSE, warning = FALSE}
url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid = read_csv(url)

setwd("/Users/Hannah/github/geog-13-labs/data")
landData = read_csv("landdata-states.csv")
popEst = read_excel("PopulationEstimates.xls", skip = 2)
```

# Question 1: Which California counties are safe?

```{r, message = FALSE, warning = FALSE}

# Create subset of daily new cases in California

daily_new = filter(covid, state %in% c("California")) %>%
group_by(county) %>%
mutate(cases_new = cases - lag(cases)) %>%
ungroup() %>%
filter(date == max(date))

# Generate table of the 5 counties with the most cumulative cases

top5_cum = arrange(daily_new, -cases) %>%
  head(n = 5) %>% 
	select(county, cases)
knitr::kable(top5_cum, caption = "CA's Top 5 Counties with the Most Cumulative Cases", col.names = c("County", "Cumulative Cases"))

# Generate table of the 5 counties with the most new cases

top5_new = arrange(daily_new, -cases_new) %>%
  head(n = 5) %>%
	select(county, cases_new)
knitr::kable(top5_new, caption = "CA's Top 5 Counties with the Most New Cases", col.names = c("County", "New Cases"))

# Join the population data with the COVID data

setwd("/Users/Hannah/github/geog-13-labs/data")
population_estimates = read_excel("PopulationEstimates.xls", skip = 2) %>% 
select(pop_2019 = POP_ESTIMATE_2019, State, fips = FIPStxt)

population_interest = population_estimates %>% filter(State == "CA")

cases_per_capita = right_join(daily_new, population_estimates, by = "fips") %>% 
  mutate(cum_cases_per_capita = (cases / pop_2019) * 100000) %>% 
  mutate(new_cases_per_capita = (cases_new / pop_2019) * 100000)

# Generate table of the 5 counties with the most cumulative cases per capita

top5_cum_per_capita = arrange(cases_per_capita, -cum_cases_per_capita) %>%
  head(n = 5) %>% 
	select(county, cum_cases_per_capita)

knitr::kable(top5_cum_per_capita, caption = "CA's Top 5 Counties with the Most Cumulative Cases Per Capita", col.names = c("County", "Cumulative Cases Per Capita"))

# Generate table of the 5 counties with the most new cases per capita

top5_new_per_capita = arrange(cases_per_capita, -new_cases_per_capita) %>%
  head(n = 5) %>% 
	select(county, new_cases_per_capita)

knitr::kable(top5_new_per_capita, caption = "CA's Top 5 Counties with the Most New Cases Per Capita", col.names = c("County", "New Cases Per Capita"))

# Finding total cases in last 14 days

cases_per_capita_recent = right_join(population_estimates, covid, by = "fips") %>%
  filter(state == "California") %>%
  filter(date >= max(date) - 13)

tot_cases = cases_per_capita_recent %>% 
  group_by(county) %>% 
  summarize(tot = sum(cases, na.rm = TRUE)) %>% 
  ungroup()

counties = cases_per_capita_recent %>%
  group_by(county, pop_2019) %>%
  mutate(cases_new = max(cases) - min(cases), cases_new_per_capita = (cases_new / pop_2019) * 100000) %>%
  ungroup()

day14_new = cases_per_capita_recent %>% 
 group_by(county, pop_2019) %>% 
 slice_max(cases, n=14) %>% 
 arrange(-cases)
```

## Overview of safe counties + new and cumulative cases:
#### As of `r as.Date(max(cases_per_capita_recent$date))`, there are `r format(max(tot_cases$tot))` cumulative coronavirus cases in CA, of which `r format(day14_new[[7]][1] - day14_new[[7]][14])` are new.
#### Per California Department of Public Health standards, there are `r length(filter(counties, cases_new_per_capita < 100)$county)` safe counties.

# Question 2: What is the impact of scale on data interpretation?

```{r, message = FALSE, warning = FALSE}

# Set up states of interest

states_of_interest <- c("New York", "California", "Louisiana", "Florida")
states_of_interest_abbr <- c("NY", "CA", "LA", "FL")

# Modify data and filter for states of interest

multi_cases = covid %>% filter(state %in% states_of_interest) %>% 
  group_by(fips) %>% 
  mutate(cases_new = cases - lag(cases)) %>% 
  ungroup() %>%
  group_by(state, date) %>%
  summarize(cases = sum(cases, na.rm = TRUE)) %>% 
  mutate(daily_new_cases = cases - lag(cases), RollMean7 = rollmean(daily_new_cases, 7, fill = NA, align = "right")) %>%
  ungroup()

# Plot the 'new cases' data

multi_cases %>%
  ggplot(aes(x = date, y = daily_new_cases)) +
  geom_line(aes(col=state)) +
  geom_line(aes(y = RollMean7), col = "grey55", size = 0.5) +
  ggthemes::theme_calc() +
  facet_wrap(~state) +
  labs(title = "Daily New Cases",
       x = "Date",
       y = "Cases",
       subtitle = "New York Times Data")

# Join the state COVID data to the USDA population estimates

int_join = covid %>%
  filter(state %in% states_of_interest) %>%
  right_join(population_estimates, by = "fips")

# Modify the data to get per capita cases

multi_cases_2 = population_estimates %>% 
  filter(State %in% states_of_interest_abbr) %>% 
  group_by(State) %>%
  summarize(pop_2019 = max(pop_2019)) %>%
  right_join(int_join, by = "State") %>% 
  ungroup() %>% 
  filter(state %in% states_of_interest) %>%
  group_by(pop_2019.x, state, date) %>% 
  summarize(cases = max(cases)) %>%
  mutate(daily_new_cases = cases - lag(cases),
         daily_new_cases_pc = daily_new_cases / pop_2019.x,
         RollMean7 = rollmean(daily_new_cases_pc, 7, fill = NA, align = "right")) %>%
  filter(RollMean7 != "NA") %>%
  ungroup()

# Plot the 'new cases per capita' data

multi_cases_2 %>%
  ggplot(aes(x = date, y = daily_new_cases_pc)) +
  geom_line(aes(col = state)) +
  geom_line(aes(y = RollMean7), col = "grey55", size = 0.5) +
  ggthemes::theme_calc() +
  facet_wrap(~state) +
  labs(title = "Daily New Cases Per Capita",
       x = "Date",
       y = "Cases Per Capita ",
       subtitle = "New York Times Data")
```

### Scaling by population has a significant effect on the analysis of this data. For example, based on the "Daily New Cases" chart alone, you might assume that New York had a much higher rate of COVID cases than Louisiana. However, the "Daily New Cases Per Capita" chart shows that Louisiana had a very similar per capita rate of COVID in relation to New York.
